Collaborative Visual Analysis on the Web with RCloud

### 2015-06-23 discussion.tex
reduced text in Versioning ("and Upgrading", as it's now called). setion 6.2 pass is done.

\paragraph*{Versioning}
Some important aspects of the environment are more difficult to manage
in RCloud than in conventional systems. RCloud does not explicitly
separate the development and deployment environments. More than that,
every version of every prototype is shared by default. Although this encourages
collaboration on work in progress it also quickly exposes errors such as
misconfiguration, mismatches between versions of packages, and programming
errors that can unintentionally affect production websites. Such errors
are difficult to completely avoid, but the power of convenient sharing is often
worth it (as proven true with distributed systems in general).
Still, effective control over versioning when sharing problem needs more
attention if RCloud would be scaled up to a wider number of users lacking
informal channels to coordinate work.

Versioning needs to be managed in several places: in scripts;
in the R environment such as the installed libraries and packages;
and in the external environment such as the operating system and
protocols spoken by remote services. At one end of the scale, we have
full control over the versioning of scripts via git, along with conventions
to name stable or working versions of scripts, and versions known to work together.
The R environment itself is under the control of its package manager
that has rules to ensure reasonable consistency.
It is at least possible for RCloud to access information about
package versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not strong.
At the other end of the scale, there is not much reason to expect version
control for the external environment. Most programmers rely on clean living
and a careful approach to system upgrades.

On top of this, RCloud allows or even encourages hackers to fork experiments
to try new ideas quickly. So far we have not done much more to address the
problem of having a lot of bits and pieces of code and data lying around,
though we have provided a framework in which it should be easier to find them
and to apply algorithms and metadata to organize them.

@@

\paragraph*{Security}
Because RCloud allows arbitrary R code to be run
\emph{by design}, it's not straightforward to protect deployments
from running unauthorized arbitrary code. RCloud uses an object capability model
\cite{Miller:2006:RCT} recently added to the Rserve
protocol~\cite{Urbanek:2003:AFW}. Object capabilities are opaque handles
to remote procedure calls; this way, web browsers never directly instruct the
RCloud backend to execute arbitrary code, preventing unauthenticated clients
from making unauthorized calls to the RCloud runtime environment.

Our back-end environment, on the other hand, assumes a high degree of trust between users:
access control for information security is delegated to the host
operating system and web server. In practice, operating
systems rely either on coarse permission models,
which tend to be ineffective, or on detailed access control
lists, which tend to be cumbersome. This means that information security
in RCloud is a delicate, unsolved matter. More
sophisticated approaches, such as information flow analysis,
and query languages that ensure differential privacy, are
still research topics~\cite{Moore:2011:SAF}.
% We are hopeful that having a richer
% environment for collaboration and information sharing will encourage
% new approaches to information security in visual analytics.
%\gordon{This sounds like we think we are helping by pushing the
%system to the breaking point, is that the idea?}

>>>>

\paragraph*{Versioning and Upgrading}
Some important aspects of the environment are more difficult to manage
in RCloud than in conventional systems. RCloud does not separate the
development and deployment environments, and every version of every
prototype is shared by default. Although this encourages
collaboration, it also quickly exposes errors such as
misconfiguration, mismatches between versions of packages, and
programming errors that can unintentionally affect production
websites. For example, if the RCloud website is distributed across
servers in different machines, different versions of system software
can cause errors that are hard to find and fix.

Versioning of different software pieces must be managed in several
levels: in data analysis and visualization scripts (that is, the
notebooks themselves); in the R environment such as the installed
libraries and packages; and in the external environment such as the
operating system, libraries, and protocols spoken by remote
services. At one end of the scale, we have full control over the
versions via git, along with conventions that fix stable notebook
versions.  The R environment itself is under the control of its
package manager that has rules to ensure reasonable consistency.  It
is at least possible for RCloud to access information about package
versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not
strong.  At the other end of the scale, there is not much reason to
expect version control for the external environment, and consider what
would happend if a web service that is used by a notebook changes the
format of its output: notebooks now stop working, but nothing in
RCloud has changed.

@@

\paragraph*{Security}
Because RCloud allows arbitrary R code to be run
\emph{by design}, it's not straightforward to protect deployments
from running unauthorized arbitrary code. RCloud uses an object capability model
\cite{Miller:2006:RCT} recently added to the Rserve
protocol~\cite{Urbanek:2003:AFW}. Object capabilities are opaque handles
to remote procedure calls; this way, web browsers never directly instruct the
RCloud backend to execute arbitrary code, preventing unauthenticated clients
from making unauthorized calls to the RCloud runtime environment.
%
Our back-end environment, on the other hand, assumes a high degree of trust between users:
access control for information security is delegated to the host
operating system and web server. In practice, operating
systems rely either on coarse permission models,
which tend to be ineffective, or on detailed access control
lists, which tend to be cumbersome. This means that information security
in RCloud is a delicate, unsolved matter. More
sophisticated approaches, such as information flow analysis,
and query languages that ensure differential privacy, are
still research topics~\cite{Moore:2011:SAF}.
% We are hopeful that having a richer
% environment for collaboration and information sharing will encourage
% new approaches to information security in visual analytics.
%\gordon{This sounds like we think we are helping by pushing the
%system to the breaking point, is that the idea?}
